---
title: "Chapyer 2: The Beta Distribution"
output: html_notebook
---
Introduction to Empriical Bayes


The beta explained on [Wolfram MathWorld](http://mathworld.wolfram.com/BetaDistribution.html) 

To summarize, the Beta Distribution has two free parameters, alpha and beta. The distribution is used as a prior distribution for binomial proportions. The domain is [0, 1] and the rang is the density of beta. It is the probibility distribution of probibilities. It represents all possible values of a probibility when we don't know what that probibility is. The mean is alpha / (alpha + beta).

The book uses data on baseball as an example and looked at batting averages. I will use data on basketball because I like that sport better.

```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

```{r}
library(readr)
nba_2016 <- read_csv("~/Emekas Documents/R-Directory/NBA 2016.csv") %>%
    mutate('Team' = as.factor(nba_2016$Team)) %>%
    mutate('PS' = as.factor(nba_2016$PS))
```

The book uses alpha = 81 and beta = 219 making the mean = .27. At the time of this writting, I don't know why those numbers were picked, but here is what I will use for my example

```{r}
## alpha
mean(nba_2016$FGM)

## alpha + beta
mean(nba_2016$FGA)

## beta
mean(nba_2016$FGA) - mean(nba_2016$FGM)

## mean
mean(nba_2016$FGM) / mean(nba_2016$FGA)



```

I took the mean of field goals made and field goals attempted. So alpha = 186, beta = 230, and the mean = .45. This isn't scientific, but it's acceptable for just learning.

Here is my beta distribution of field goal percentatge
```{r}
data.frame(x = seq(0, 1, .01), y = dbeta(seq(0, 1, .01), 187, 230)) %>%
  ggplot(., aes(x = x)) +
  geom_line(aes(y = y)) +
  xlim(0, 1) +
  xlab('Field Goal Percentage') +
  ylab('Density of Beta') 
```

The beta distribtion is appropriate for modeling the biomial distribtion. We start with a prior distribution, obtain evidence, then update it to a posterior distribution. The curve changes each time someone takes a shot.

Lets say for example, a player has a good game and they go 15/20 from the field. Now alpha = (187 + 15) and beta = (230 + 5). So this is the updated beta distribution. 


```{r}
data.frame(x = seq(0, 1, .01), 
           y = dbeta(seq(0, 1, .01), 187, 230), 
           z = dbeta(seq(0, 1, .01), (187+15), (230 + 5))) %>%
  gather(., -x, key = func, value = val) %>%
  ggplot(., aes(x = x, y = val, group = func)) +
    geom_line(aes(color = func), size = .8) + 
    scale_color_manual(name="Beta params", 
                     values=c("black", "red"),
                     labels=c("??=187, ??=230", "??=202, ??=235")) +
    labs(x="Field Goal Percentage", y="Density of Beta")
```

So now there is a new range of possible values for that player's possible true shooting percentage. The posterior mean equals

```{r}
## (alpha + makes) / ((alpha + makes) + (beta + misses))
(187+15) / ((187+15) + (230 + 5))
```
So this good shooting performance increases this players range of possible true shooting percentages. However, it would be silly to think a player actually shoots 75% from the field based on one game.




