---
title: "Chapter 8: Empirical Bayesian Hierarchical Modeling"
output: html_notebook
---
The chapter proposes a situation in which you are choosing between two batters who each have 30/100 hits. The only difference is one is left-handed and the other is right-handed. Right handed batters tend to hit better than left-handed batters. So in this chapter we are using additional factors to influence our prior distribution.


##8.1 Setup.
Same ish
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
NBA_2016 <- read_csv("~/Emekas Documents/R-Directory/NBA 2016.csv")
```

```{r}
shooting <- NBA_2016 %>%
  filter(FGA > 0) %>%
  mutate(FGN = FGA - FGM, average = FGM/FGA) %>%
  dplyr::select(Player, Team, PS, FGM, FGN, FGA, average)

# from chapter 3
alpha0 <- 39.23353 
beta0 <- 48.72738
prior_mu <- (alpha0) / (alpha0 + beta0)

shooting_eb <- shooting %>%
  mutate(eb_estimate = ((FGM + alpha0) / (FGA + alpha0 + beta0)),
         alpha1 = alpha0 + FGM, 
         beta1 = beta0 + FGN)

shooting_eb <- shooting_eb %>%
  mutate(alpha1 = alpha0 + FGM, beta1 = beta0 + FGN)
```

from chapter 7 we will perform beta binomial regression using the gamlss package so mu depends on the number of shots taken.
```{r}
library(gamlss)

fit <- gamlss(cbind(FGM, FGA - FGM) ~ log(FGA),
              data = dplyr::select(shooting, -PS),
              family = BB(mu.link = "identity"))
```

alpha0 and beta0 can be computed based on mu and sigma.
```{r}
shooting_eb <- shooting %>%
  mutate(mu = fitted(fit, "mu"),
         sigma = fitted(fit, "sigma"),
         alpha0 = mu / sigma,
         beta0 = (1 - mu) / sigma,
         alpha1 = alpha0 + FGM,
         beta1 = beta0 + FGA - FGM,
         estimate = alpha1 / (alpha1 + beta1))
```

Noow we have corrected for the confounding factor of FGA.

##8.2 Right-handed and Left-handed Batters
Left handed batters tend to hit better than right-handed batters. Well, I'm using a basketball data set so I will use position (point guard, shooting guard, small foward, power foward, center) instead. position has a huge affect on FG% because centers like Deandre Jordan mostly shoot high percentage shots. Hopefully, this still works out.
```{r}
shooting %>%
  count(PS)
```

relevel the factors so point guards are the baseline.
```{r}
shooting2 <- shooting
shooting2$PS <- factor(shooting2$PS, levels = c("PG", "SG", "SF", "PF", "C"))

fit2 <- gamlss(cbind(FGM, FGA - FGM) ~ log(FGA) + PS,
              data = shooting2,
              family = BB(mu.link = "identity"))
```

Now we look at the coeeficients
```{r}
library(broom)
tidy(fit2)
```

Different positions have statistically significant FG%s.


```{r}
sigma <- fitted(fit2, "sigma")[1]

crossing(PS = c("PG", "SG", "SF", "PF", "C"),
         FGA = c(1, 10, 100, 1000, 1500)) %>%
  augment(fit2, newdata = .) %>%
  rename(mu = .fitted) %>%
  crossing(x = seq(.1, .8, .0005)) %>%
  mutate(alpha = mu / sigma,
         beta = (1 - mu) / sigma,
         density = dbeta(x, alpha, beta)) %>%
  ggplot(aes(x, density, color = factor(FGA), lty = PS)) +
  geom_line() +
  labs(x = "Field Goal Percentage",
       y = "Prior density",
       color = "FGA",
       lty = "Position")
```
So this doesn't really work too well with 5 different positions. So I'm going to redo this whole thing separating Guards and Forwards/Centers

```{r}
shooting3 <- shooting2
shooting3$PS <- gsub("PG", "Guard", shooting3$PS)
shooting3$PS <- gsub("SG", "Guard", shooting3$PS)
shooting3$PS <- gsub("SF", "Forward", shooting3$PS)
shooting3$PS <- gsub("PF", "Forward", shooting3$PS)
shooting3$PS <- gsub("C", "Forward", shooting3$PS)
shooting3$PS <- as.factor(shooting3$PS)


shooting3 <- shooting3 %>%
  mutate(PS = relevel(PS, "Guard"))


fit2 <- gamlss(cbind(FGM, FGA - FGM) ~ log(FGA) + PS,
              data = shooting3,
              family = BB(mu.link = "identity"))
```

```{r}
library(broom)
tidy(fit2)
```

So forwards make shots about 5% more often than guards. It's a statistically significant difference.

```{r}
sigma <- fitted(fit2, "sigma")[1]

crossing(PS = c("Guard", "Forward"),
         FGA = c(1, 10, 100, 1000, 1500)) %>%
  augment(fit2, newdata = .) %>%
  rename(mu = .fitted) %>%
  crossing(x = seq(.1, .8, .0005)) %>%
  mutate(alpha = mu / sigma,
         beta = (1 - mu) / sigma,
         density = dbeta(x, alpha, beta)) %>%
  ggplot(aes(x, density, color = factor(FGA), lty = PS)) +
  geom_line() +
  labs(x = "Field Goal Percentage",
       y = "Prior density",
       color = "FGA",
       lty = "Position")
```

Notice that yo ucan still overcome the prior of position. The guard disributions still overlap the forward distributions. Even a guard who has only taken 1 shot can still have a higher true field goal percentage than a forward who has taken 1500 shots.

The rest has to do with historical data which doesn't work with my data so read the book for that part.

